---
title: "Marketing Dashboard"
description: "Interactive marketing performance dashboard — Q1 2025 campaign overview."
toc: false
---

```{r}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("plotly")) install.packages("plotly")
if (!require("dplyr")) install.packages("dplyr")
if (!require("knitr")) install.packages("knitr")

library(tidyverse)
library(ggplot2)
library(plotly)
library(dplyr)
library(knitr)
```


[Live Analytics]{.section-label}

<hr class="divider-gold">

This dashboard tracks key marketing KPIs across all active channels. Built using Tableau Public and R, it aggregates data from Google Analytics 4, paid social exports, and the CRM.

------------------------------------------------------------------------

## KPI Summary — Q1 2025

```{=html}
<div class="kpi-grid">
  <div class="kpi-card">
    <div class="kpi-label">Total Reach</div>
    <div class="kpi-value">142K</div>
    <div class="kpi-delta">▲ 18% vs Q4 2024</div>
  </div>
  <div class="kpi-card">
    <div class="kpi-label">Conversions</div>
    <div class="kpi-value">3,841</div>
    <div class="kpi-delta">▲ 11% vs Q4 2024</div>
  </div>
  <div class="kpi-card">
    <div class="kpi-label">Email Open Rate</div>
    <div class="kpi-value">28.4%</div>
    <div class="kpi-delta neg">▼ 2% vs Q4 2024</div>
  </div>
  <div class="kpi-card">
    <div class="kpi-label">Cost Per Lead</div>
    <div class="kpi-value">$14.20</div>
    <div class="kpi-delta">▲ Improved 9%</div>
  </div>
  <div class="kpi-card">
    <div class="kpi-label">Organic Sessions</div>
    <div class="kpi-value">68.4K</div>
    <div class="kpi-delta">▲ 34% vs Q4 2024</div>
  </div>
  <div class="kpi-card">
    <div class="kpi-label">Avg. Session Duration</div>
    <div class="kpi-value">2m 18s</div>
    <div class="kpi-delta">▲ 8% vs Q4 2024</div>
  </div>
</div>
```

------------------------------------------------------------------------


## Monthly Conversions — R Chart

```{r}
#| label: monthly-conversions
#| fig-alt: "Bar chart showing monthly conversions from October 2024 through March 2025"
#| fig-cap: "Monthly conversions across all channels, Q1 2025. February spike driven by organic search content refresh."

library(tidyverse)

conversions <- tibble(
  month = factor(
    c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar"),
    levels = c("Oct", "Nov", "Dec", "Jan", "Feb", "Mar")
  ),
  value    = c(210, 290, 340, 280, 410, 380),
  highlight = c(FALSE, FALSE, FALSE, FALSE, TRUE, FALSE)
)

ggplot(conversions, aes(x = month, y = value, fill = highlight)) +
  geom_col(width = 0.65) +
  geom_text(aes(label = scales::comma(value)), vjust = -0.5,
            size = 3.2, family = "sans", color = "#0f2444") +
  scale_fill_manual(values = c(`FALSE` = "#0f2444", `TRUE` = "#c9a84c"),
                    guide = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12)),
                     labels = scales::comma) +
  labs(x = NULL, y = "Conversions",
       caption = "Gold bar = peak month (February)") +
  theme_minimal(base_family = "sans") +
  theme(
    axis.text  = element_text(color = "#5a5a5a", size = 10),
    axis.title = element_text(color = "#0f2444", size = 10),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    plot.caption = element_text(color = "#888", size = 8)
  )
```

------------------------------------------------------------------------

## Channel Mix — R Pie / Donut

```{r}
#| label: channel-mix
#| fig-alt: "Donut chart showing marketing channel mix for Q1 2025"
#| fig-cap: "Q1 2025 channel mix by conversion volume. Organic search is the top driver."

channels <- tibble(
  channel = c("Organic Search", "Paid Social", "Email", "Direct"),
  pct     = c(42, 28, 18, 12)
) |>
  mutate(
    channel = fct_reorder(channel, pct, .desc = TRUE),
    label   = paste0(channel, "\n", pct, "%")
  )

ggplot(channels, aes(x = 2, y = pct, fill = channel)) +
  geom_col(width = 1, color = "white", linewidth = 0.8) +
  coord_polar(theta = "y", start = 0) +
  xlim(0.5, 2.5) +
  scale_fill_manual(values = c(
    "Organic Search" = "#0f2444",
    "Paid Social"    = "#c9a84c",
    "Email"          = "#1a3a6b",
    "Direct"         = "#b0a080"
  )) +
  labs(fill = "Channel") +
  theme_void(base_family = "sans") +
  theme(legend.position = "right",
        legend.text = element_text(size = 9, color = "#1c1c1c"))
```

::: {.callout-note appearance="simple"}
**Key Insight:** Organic search conversions grew 34% in February following the content refresh initiative — the strongest ROI of any channel this quarter. Recommend increasing SEO investment in Q2.
:::
