---
title: "RPE Project In-Class Activities"
subtitle: "Exploratory Data Analysis and Understanding Data"
author: "Maurice Morgan"
institute: "IBM 3202 · Cal Poly Pomona"
date: 2026-04-22
format:
  revealjs:
    theme: default
    footer: "EDA and Data Wrangling · IBM 3202"
    transition: slide
    transition-speed: fast
    toc: true
    toc-depth: 1
    toc-title: "Table of Contents"
    slide-number: true
    scrollable: true
    smaller: true
    code-fold: false
    code-overflow: wrap
    code-copy: hover
    embed-resources: true
    self-contained-math: true
    title-slide-attributes:
      data-background-color: "#0f2444"
execute:
  echo: true
  freeze: auto
  error: true
  warning: false
  message: false
---

# Introduction

## About This Presentation

This presentation walks through the full EDA pipeline applied to the RPE (Rate of Perceived Exertion) combined survey dataset.

**Goals:**

-   Load and inspect a labelled SPSS dataset
-   Apply data cleaning and standardization
-   Perform exploratory data analysis
-   Generate visualizations and initial insights

## Loading Libraries and Data {#loading}

```{r}
#| label: package-loading

# Uncomment to install if needed:
# install.packages(c("tidyverse","haven","labelled","janitor","gt","sjPlot"))

library(tidyverse)
library(haven)
library(labelled)
library(janitor)
library(gt)
library(sjPlot)
```

## Reading and Printing Data

```{r}
#| label: data-reading
#| eval: false   # set eval: true once data file is in /data/

data <- read_sav("../data/RPE-Combined-2019-12-19.sav")
data <- data |> clean_names()
data
```

::: callout-tip
**`clean_names()`** from `janitor` converts all column names to `snake_case`, removing spaces and special characters — essential for consistent downstream coding.
:::

# Data Exploration

## Data Structure and Summary

```{r}
#| label: data-structure
#| eval: false

str(data)
summary(data)
glimpse(data)

# Data dictionary — shows SPSS variable labels
view_df(data)
```

## What `glimpse()` Tells Us

`glimpse()` gives a quick transposed view of the dataset — perfect for a first look when you have many columns.

| Output | What to look for |
|---------------------|--------------------------------------------------|
| Row / column count | Dataset size, potential performance issues |
| Column types | `<dbl>`, `<chr>`, `<fct>`, `<lbl>` (labelled) |
| First values | Obvious encoding issues, unexpected values |

: {tbl-colwidths="[30,70]"}

## The `labelled` Package

SPSS `.sav` files store **variable labels** and **value labels** that R's `read_sav()` preserves as `<labelled>` vectors.

```{r}
#| label: labelled-demo
#| eval: false

# Inspect labels on one variable
data |> pull(rpe_score) |> val_labels()   # replace with actual var name

# Convert labelled vars to factors for plotting
data_clean <- data |>
  mutate(across(where(is.labelled), as_factor))
```

::: callout-note
Always decide early whether to work with labelled vectors or convert to factors — mixing both leads to downstream issues.
:::

# Exploratory Data Analysis

## EDA Overview

::: callout-tip
In this section we apply EDA techniques to:

-   **Summarize** key characteristics (mean, median, spread)
-   **Uncover** patterns and trends
-   **Detect** anomalies and outliers
-   **Assess** data quality (missing values, encoding issues)
-   **Generate** insights that guide modeling and decisions
:::

## Descriptive Statistics

```{r}
#| label: descriptive-stats
#| eval: false

# Frequency tables for categorical variables
data |>
  count(gender, sort = TRUE) |>   # replace with actual variable name
  mutate(pct = n / sum(n) * 100)

# Summary statistics for numeric variables
data |>
  summarise(across(where(is.numeric),
    list(mean = mean, sd = sd, median = median, min = min, max = max),
    na.rm = TRUE
  ))
```

## Distribution Visualization

```{r}
#| label: viz-distribution
#| eval: false
#| fig-cap: "Distribution of RPE scores — note bimodal pattern"

data |>
  ggplot(aes(x = rpe_score)) +     # replace with actual variable name
  geom_histogram(fill = "#0f2444", color = "#c9a84c",
                 bins = 20, alpha = 0.9) +
  labs(title = "Distribution of RPE Scores",
       x = "RPE Score", y = "Count") +
  theme_minimal(base_family = "sans")
```

## Correlation Analysis

```{r}
#| label: correlation
#| eval: false

library(corrplot)

data |>
  select(where(is.numeric)) |>
  cor(use = "complete.obs") |>
  corrplot(method = "color", type = "upper",
           tl.cex = 0.7, col = colorRampPalette(c("#c9a84c","white","#0f2444"))(200))
```

## Missing Data Assessment

```{r}
#| label: missing-data
#| eval: false

# Count and percentage of missings per variable
data |>
  summarise(across(everything(),
    list(n_miss = ~sum(is.na(.)),
         pct_miss = ~round(mean(is.na(.)) * 100, 1))
  )) |>
  pivot_longer(everything(),
    names_to  = c("variable", ".value"),
    names_sep = "_(?=[^_]+$)"
  ) |>
  filter(n_miss > 0) |>
  arrange(desc(pct_miss))
```

# Findings and Next Steps

## Key Takeaways

::: columns
::: {.column width="50%"}
**Data Quality**

-   X% of records complete (no missing values)
-   Y variables required re-encoding from labelled format
-   Z outliers flagged for review
:::

::: {.column width="50%"}
**Substantive Patterns**

-   RPE scores show a bimodal distribution — possible sub-population
-   Several Likert items exhibit ceiling effects (\>70% at max)
-   Strong correlation between \[var A\] and \[var B\] (*r* = 0.XX)
:::
:::

## Recommended Next Steps

1.  **Deeper missing data analysis** — investigate whether missingness is random (MAR) or systematic (MNAR)
2.  **Sub-group analysis** — split by demographic variable to explore bimodal RPE pattern
3.  **Feature engineering** — create composite scores from correlated Likert items
4.  **Predictive modeling** — logistic or OLS regression on key outcome variable

## References and Resources

| Resource | Link |
|-------------------------|-----------------------------------------------|
| tidyverse documentation | <https://tidyverse.org> |
| haven package | <https://haven.tidyverse.org> |
| labelled package | <https://larmarange.github.io/labelled> |
| janitor package | <https://sfirke.github.io/janitor> |
| sjPlot package | <https://strengejacke.github.io/sjPlot> |

: {tbl-colwidths="[35,65]"}
