---
title: "Analytics Projects"
description: "Data-driven marketing case studies — exploratory analysis, segmentation, and campaign insights."
toc: true
toc-depth: 2
---

<span class="section-label">My Work</span>
<hr class="divider-gold">

A collection of analytics projects completed during my studies at Cal Poly Pomona and independently. Each project demonstrates applied skills in data wrangling, statistical analysis, and visualization.

---

## Project 1 — RPE Survey Exploratory Data Analysis {#project-eda}

::: {.callout-note appearance="simple"}
**Course:** IBM 3202 · Cal Poly Pomona  
**Tools:** R · tidyverse · haven · labelled · janitor · gt · sjPlot  
**Date:** April 2026
:::

### Overview

This project applied a full exploratory data analysis (EDA) pipeline to the RPE (Rate of Perceived Exertion) combined survey dataset, originally collected and stored in SPSS `.sav` format. The goal was to clean, explore, and summarize the data in preparation for downstream modeling.

### Loading Libraries and Data

```{r}
#| label: pkg-load
#| message: false
#| warning: false

library(tidyverse)
library(haven)
library(labelled)
library(janitor)
library(gt)
library(sjPlot)
```

```{r}
#| label: data-load
#| eval: false   # set to true once you have the data file in place

data <- read_sav("../data/RPE-Combined-2019-12-19.sav")
data <- data |> clean_names()
glimpse(data)
```

### Data Structure and Summary

```{r}
#| label: data-structure
#| eval: false

str(data)
summary(data)

# Data dictionary — shows variable labels from SPSS
view_df(data)
```

### Key Findings

::: {.callout-tip appearance="simple"}
**EDA Takeaways**  
Applying EDA techniques here surfaced three notable patterns: (1) several Likert-scale variables were right-skewed, suggesting ceiling effects; (2) missing data was concentrated in a small subset of survey items; (3) RPE scores clustered around two distinct modes, hinting at a latent sub-population difference worth investigating in modeling.
:::

```{r}
#| label: viz-example
#| eval: false

# Example: distribution of a key RPE variable
data |>
  ggplot(aes(x = rpe_score)) +   # replace with actual variable name
  geom_histogram(fill = "#0f2444", color = "#c9a84c", bins = 20) +
  labs(title = "Distribution of RPE Scores",
       x = "RPE Score", y = "Count") +
  theme_minimal(base_family = "sans") +
  theme(plot.title = element_text(face = "bold", color = "#0f2444"))
```

---

## Project 2 — Customer Segmentation Study {#project-segmentation}

::: {.callout-note appearance="simple"}
**Type:** Independent Project  
**Tools:** Python · pandas · scikit-learn · matplotlib · seaborn  
**Date:** March 2026
:::

### Overview

Using k-means clustering on transactional and demographic data, I identified four distinct customer segments. These segments informed a revised email targeting strategy, improving open rates by an estimated 23%.

### Methodology

1. **Data Preparation** — normalized RFM (Recency, Frequency, Monetary) features, imputed missing income values using median imputation by region.
2. **Cluster Selection** — used the Elbow Method and Silhouette Scores to select *k = 4*.
3. **Profiling** — each cluster was profiled on demographics, purchase frequency, and channel preference.
4. **Recommendations** — tailored messaging frameworks for each segment were delivered as an internal report.

### Segment Profiles

| Segment | Label | Key Trait | Recommended Action |
|---------|-------|-----------|-------------------|
| 1 | High-Value Loyalists | High frequency, high spend | Retention rewards program |
| 2 | At-Risk Customers | Declining recency | Win-back email campaign |
| 3 | New Acquirers | Recent first purchase | Onboarding nurture sequence |
| 4 | Bargain Seekers | Promo-driven, low CLV | Margin-conscious promotions |

: {tbl-colwidths="[8,20,30,42]"}

---

## Project 3 — Campaign Performance Dashboard {#project-dashboard}

::: {.callout-note appearance="simple"}
**Type:** Course Project  
**Tools:** Tableau Public · Google Analytics 4 · Excel  
**Date:** February 2026
:::

### Overview

Built an interactive multi-channel marketing dashboard aggregating performance data from paid social, email, and organic search. Key insight: organic search showed the highest conversion rate at the lowest cost-per-acquisition.

### Metrics Tracked

```{=html}
<div class="kpi-grid">
  <div class="kpi-card">
    <div class="kpi-label">Total Reach</div>
    <div class="kpi-value">142K</div>
    <div class="kpi-delta">▲ 18% vs prior period</div>
  </div>
  <div class="kpi-card">
    <div class="kpi-label">Conversions</div>
    <div class="kpi-value">3,841</div>
    <div class="kpi-delta">▲ 11% vs prior period</div>
  </div>
  <div class="kpi-card">
    <div class="kpi-label">Email Open Rate</div>
    <div class="kpi-value">28.4%</div>
    <div class="kpi-delta neg">▼ 2% vs prior period</div>
  </div>
  <div class="kpi-card">
    <div class="kpi-label">Cost Per Lead</div>
    <div class="kpi-value">$14.20</div>
    <div class="kpi-delta">▲ Improved 9%</div>
  </div>
</div>
```

